---
title: "Proteomics cell lines"
output: html_notebook
---

### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/Proteomics_virus/")) 
```

```{r}
library(ggplot2)
library(limma)
library(dplyr)
library(xlsx)
library(RColorBrewer)
library(gplots)
library(ggrepel)
library(knitr)
library(ggfortify)
library(NormalyzerDE)
library(ropls)
library(stringr)
opts_chunk$set(echo = TRUE)
options(bitmapType="cairo")
```

```{r}
source("src/usefull_functions.R")
```

```{r}
input_1 <- "data/Cell_lines/293T_Huh7_TMTpro_12Fractions.csv"
input_2 <- "data/Cell_lines/Calu2_Caco3_TMTpro_12Fractions_2.csv"
n <- 3
```



## 1) Pre-processing

## file 1

```{r}
name_ana <- "Calu3"
```
```{r}
dir.create(paste0("processing/", name_ana))
dir.create(paste0("results/", name_ana))
dir.create(paste0("results/", name_ana, "/figures"))
dir.create(paste0("results/", name_ana, "/GSEA"))
dir.create(paste0("results/", name_ana, "/DGE"))
```

```{r}
data <- read.csv(input_2)
data$Description <- NULL
data <- data[,1:7]
samples <- data.frame(names = colnames(data)[-1], condition = NA, names_2 = NA)
samples$condition <- c("R2_Calu3_U","R2_Calu3_T", "R2_Calu3_U","R2_Calu3_T","R2_Calu3_U","R2_Calu3_T")
samples$names_2 <- paste0(samples$condition, c(rep(1,2), rep(2,2), rep(3,2))) ## extract run number
samples

condition <- samples$condition

colnames(data)[-1] <- samples$names_2
rownames(data) <- data$Accession
data$Accession <- NULL
```

## Make annotation_info_file
```{r}
annotationInfo <- read.delim("data/annotationInfo_Uniprot_gene_name")
names(annotationInfo)[1] <- "Accession"
```

## select Abundances columns
```{r}
print(nrow(data))
data <- data[rowSums(is.na(data)) != ncol(data), ] ## remove empty lines
data_trash <- data[!complete.cases(data),]
data <- data[complete.cases(data),]
print(nrow(data))
```
## save files
```{r}
save_file_csv(data, "row_data_filt", paste0("processing/", name_ana, "/"))

save_file_csv(samples, "informations", paste0("processing/", name_ana, "/"))
```

### check for missing values
```{r}
counts <- 1:nrow(data) # create vector of appropriate length
for(i in 1:nrow(data)){
    # TRUE will be coerced to 1 for the summing
    counts[i] <- sum(is.na(data[i,]))
}
table(counts) # create the summary
```
### distribution
```{r}
data_2 <-as.matrix(data)
hist(data_2 , prob=T)
curve(dnorm(x,mean(data_2),sd(data_2)),add=T,col="red")
save_figure(data_2, "distribution_", 10, 10)
```


### PCA before norm
```{r}
pca_data <- data.frame(group = condition, 
                       t(as.matrix(data)))
```

```{r}
pca_data[is.na(pca_data)] <-  0
pca <- prcomp(pca_data[,-1])
```
+  scale_fill_manual(values = c(c_c2, c_c1))
```{r}
a <- autoplot(pca, data = pca_data, size= 3, shape = 21, fill = 'group')+theme_gray()
print(a)
save_figure(a, "pca_before_norm_", 4, 4)
```


### Quantile normalization

```{r}
path_data <- make_data_matrix(data, name_ana)
path_design <-make_design_matrix(data, name_ana, condition)
outDir <- "processing/normalization"
```

```{r}
#normalyzer(jobName=name_ana, designPath=path_design, dataPath=path_data, outputDir=outDir)
```
```{r}
path_results_norm <- paste0(outDir, "/", name_ana,"/Quantile-normalized.txt")
data_norm <- read.delim(path_results_norm)
rownames(data_norm) <- rownames(data)

path_results <- paste0("processing/norm_files/", name_ana, "_norm_file.txt")

write.table(data_norm, file = path_results, 
            sep = "\t", row.names = TRUE, col.names = NA)
```

### PCA after normalization
```{r}
pca_data <- data.frame(group = condition, t(as.matrix(data_norm)))
```

```{r}
pca_data[is.na(pca_data)] <-  0
pca <- prcomp(pca_data[,-1])
```
+ scale_fill_manual(values = c(c_c2, c_c1))
```{r}
b <- autoplot(pca, data = pca_data, size= 3, shape = 21, fill = 'group')+theme_gray()
print(b)
save_figure(b, "pca_after_norm_", 4, 4)
```

```{r}
#c <- autoplot(pca, data = pca_data, size= 6, fill = 'group', label = TRUE, label.size = 3, shape = FALSE)+theme_gray()
#print(c)
#save_figure(c, "pca_after_norm_label", 4, 4)
```

### distribution after normalization
```{r}
#data_2 <-as.matrix(data_norm)
#hist(data_2 , prob=T)
#curve(dnorm(x,mean(data_2),sd(data_2)),add=T,col="red")
#save_figure(data_2, "distribution_", 10, 10)
```
## Multiple comparisons

## four time points experiments = 

## set levels
```{r}
groups <- as.factor(condition)
design <- model.matrix( ~ 0 + groups)
colnames(design) <- levels(groups)
```

## fit model
```{r}
fit <- lmFit(data_norm, design)
```


```{r}
list_comp <- c("R2_Calu3_T-R2_Calu3_U")
```

```{r}
head(data_norm)
```
```{r}
viral_proteins <- c("P0DTC2", "P0DTC3", "P0DTC5", "P0DTC6", "P0DTC7", "P0DTC8", "P0DTC9", "P0DTD1", "P0DTD2")
```

```{r}
length(viral_proteins[viral_proteins %in% rownames(data_norm)])
```
```{r}
data_virus <- data_norm[rownames(data_norm) %in% viral_proteins,]
```

```{r}
data_virus_2 <- data.frame(Proteins = rownames(data_virus),
                           S1 = data_virus[,2] - data_virus[,1],
                           S2 = data_virus[,4] - data_virus[,3],
                           S3 = data_virus[,6] - data_virus[,5])
```
```{r}
library(reshape2)
data_virus_2 <- melt(data_virus_2, by = "Proteins")
```
```{r}
pdf("results/virus_levels/Calu3_virus_levels_difference.pdf")
a<- ggplot(data_virus_2, aes(x=Proteins, y=value, fill = "#1F78B4"))+
  geom_bar(stat="identity", position=position_dodge())+
  scale_fill_manual(values =c("#1F78B4"))+ theme_minimal()+ 
  theme(legend.position = "none", legend.title=element_text(size=8),legend.text=element_text(size=8),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=15, color = "black"),axis.title.x=element_text(size=10, color = "black"),
        axis.text.y=element_text(size=12, color = "black"),axis.text.x=element_text(size=10, color = "black"))
a
dev.off()
```
```{r}
data_virus_2$Cell_line <- name_ana
write.csv(data_virus_2, paste0("results/virus_levels/", name_ana, "differences_viral_proteins.csv"))
```

```{r}

data_virus <- data.frame(condition = condition, t(data_virus))
data_virus <- melt(data_virus, by = "condition")
```
```{r}
#+++++++++++++++++++++++++
# Function to calculate the mean and the standard deviation
  # for each group
#+++++++++++++++++++++++++
# data : a data frame
# varname : the name of a column containing the variable
  #to be summariezed
# groupnames : vector of column names to be used as
  # grouping variables
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}
```

```{r}
data_virus <- data_summary(data_virus, varname="value", 
                    groupnames=c("condition", "variable"))
# Convert dose to a factor variable
data_virus$variable =as.factor(data_virus$variable)
head(data_virus)
```
```{r}
brewer.pal(n = 8, name = "Paired")
```

```{r}
pdf("results/virus_levels/Calu3_virus_levels.pdf", height = 5)
a<- ggplot(data_virus, aes(x=variable, y=value, fill = condition))+
  geom_bar(stat="identity", position=position_dodge())+
  scale_fill_manual(values =c("#1F78B4","#A6CEE3"))+ 
  theme(legend.title=element_text(size=8),legend.text=element_text(size=12),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=15, color = "black"),axis.title.x=element_text(size=10, color = "black"),
        axis.text.y=element_text(size=12, color = "black"),axis.text.x=element_text(size=10, color = "black"))+
  geom_errorbar(aes(ymin=value-sd, ymax=value+sd), width=.2,
                 position=position_dodge(.9))+ theme_minimal()
a
dev.off()
```


```{r}
#theme_set(theme_classic())
a<- ggplot(table, aes(x=Term, y=-log10(Adjusted.P.value), fill = Sign))+
  geom_bar(stat="identity", colour="black", size = 0.3)+
  scale_x_discrete(limits = position) + coord_flip()+
  scale_fill_manual(values =c("#009933","#cc0000")) + geom_text(aes(label=Overlap), size=3, color = "white", hjust = 1.1)+ 
  theme(legend.title=element_text(size=8),legend.text=element_text(size=6),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=15, color = "black"),axis.title.x=element_text(size=15, color = "black"),
        axis.text.y=element_text(size=12, color = "black"),axis.text.x=element_text(size=12, color = "black"))
 
a
```


```{r}
for(comp in list_comp){
  c2 <- substr(comp, 1,10)[1]
  c1 <- substr(comp, 12,21)[1]
  dge <- extract_limma(fit, design, name_ana, c1, c2, viral_proteins)
  print(dge)
  dat <- select(data_norm, contains(c1), contains(c2))
  clustering_DGE(dat, dge, name_ana, comp, annotationInfo)
  top_table_up <- dge[dge$Significance_2 == "upregulated",]
  top_table_down <- dge[dge$Significance_2 == "downregulated",]
  print(top_table_up)
  name_up <- paste0("results/",name_ana,"/GSEA/prot_up_regulated_", comp, ".txt")
  name_down <- paste0("results/",name_ana,"/GSEA/prot_down_regulated_", comp, ".txt")
  print(name_up)
  print(name_down)
  write.table(data.frame(top_table_up$Gene_name), file = name_up, sep = "\t",
            row.names = FALSE)
  write.table(data.frame(data = top_table_down$Gene_name), file = name_down, sep = "\t",
            row.names = FALSE)
}
```